Applying Multi-Core Model Checking to Hardware-Software Partitioning in Embedded Systems

### 2015-08-23 Docs/Paper/LaTex/template.tex
Update Reviews Sec 2.2/3.2/8/9

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestore
\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[twocolumn]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{color}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{booktabs}
\usepackage{tabu,xcolor,colortbl}
\usepackage{rotating}
\usepackage{amsmath,array}    
\usepackage[utf8]{inputenc} 
\DeclareUnicodeCharacter{FB01}{fi} %Caracter especial
\usepackage{graphicx} %Imagens
\usepackage{listings} % Codes
\usepackage{tabularx} % in the preamble
\usepackage{mathtools} % Serve para as equacoes

\usepackage{hyphenat}
\usepackage[english]{babel}
\hyphenation{ef-fect-ive consi-de-red demons-trate se-cond veri-fication covera-ge des-cribe reali-zation opera-tor re-pre-sentation methodo-lo-gy di-gi-tal chara-cteristics granu-lar ge-nerally diago-nal para-meters spe-cification re-presentable me-thods ope-rations sche-dule deve-loped des-cribes diffe-rent mo-dels sa-ving ve-rified proper-ties mini-mum mathe-matical tole-rated limi-ted op-tical net-works semi-conduc-tor go-to-pro-grams}

@@

\begin{document}
\title{Applying Multi-Core Model Checking to Hardware-Software Partitioning in Embedded Systems}
\author{Alessandro Trindade\and Hussama Ismail\and Renato Degelo\and Edilson Galv\~ao\and Lucas Cordeiro}

\institute{A. Trindade, H. Ismail, R. Degelo, E. Galv\~ao \at
              Graduate Program in Electrical Engineering, Federal University of Amazonas, Brazil \\
%              Tel.: +123-45-678910\\
%              Fax: +123-45-678910\\
              \email{\{iurybessa, hussamaibrahim\}@ufam.edu.br and\\{\{rdegelo, esj.galvao\}@gmail.com}}          %  \\
%             \emph{Present address:} of F. Author  %  if needed
           \and
           L. Cordeiro \at
              Electronic and Information Research Center, Federal University of Amazonas, Brazil
              \email: lucascordeiro@ufam.edu.br           %  \\
}

@@

\begin{abstract}
\textcolor{blue}{Atualizar o abstract para descrever brevemente os novos algoritmos desenvolvidos}
We present an alternative approach to solve the hardware (HW) and software (SW) partitioning problem, which uses Bounded Model Checking (BMC) based on Satisfiability Modulo Theories (SMT) in conjunction with a multi-core support using Open MultiProcessing. The multi-core SMT-based BMC approach allows initializing many verification instances based on processors cores numbers available to the model checker. Each instance checks for a different optimum value until the optimization problem is satisfied. The goal is to show that multi-core model-checking techniques can be effective, in particular cases, to find the optimal solution of the HW-SW partitioning problem using an SMT-based BMC approach. \textcolor{Red}{The probem of HW-Sw partitioning was modelled on many platforms such as: MATLAB R2013a by MathWorks[19], Z3 a theorem solver by Microsoft Research[32] and ESBMC using single-core and multi-core approach. The implemented algorithms are discussed and it's main features highlighted, finally, a comparison of the results is made. }
\keywords{hardware-software co-design \and hardware-software partitioning\and optimization\and model checking\and multi-core\and OpenMP }
\end{abstract}

\section{Introduction}
Nowadays, with the strong development of embedded systems, the design phase plays an important role. At early stages, the design is split into separated flows: hardware and software. Consequently, the partitioning decision process, which deals with the decisions upon which parts of the application have to be designed in hardware (HW) and which in software (SW), must be supported by any well-structured methodology. If not, this leads to a number of issues (design flow interruptions, redesigns, and undesired iterations) which affects the overall development process, the quality and the lifecycle of the final system. Starting at the 1990s, intensive research was performed, and several approaches proposed, as shown in [1] and [2].

In any HW and SW design of complex systems, more time is spent on verification than on construction [3]. Formal methods based on model checking offer great potential to obtain a more effective and faster verification in the design process. Programs may be viewed as mathematical objects with behavior that is, in principle, well determined. This makes it possible to specify programs using mathematical logic, which constitutes the intended (correct) behavior. Then, one can try to give a formal proof or otherwise establish that the program meets its specification [4]. Research in formal methods has led to the development of very promising verification techniques, which facilitate the early detection of errors. Model-based verification techniques use models that describe the possible system behavior in a mathematically precise and unambiguous manner. The system models are accompanied by algorithms that systematically explore all the states of the system model.

In [5] and [6] was shown that it is possible to use Bounded Model Checking (BMC) based on Satisfiability Modulo Theories (SMT) to perform HW-SW partitioning in embedded systems. The present work extends those studies since there is a substantial improvement in terms of the genetic algorithm and the SMT-based verification method, which has been extended with a multi-core architecture. Multi-core processors have been used in all segments of industry to implement high-performance computing [7]. In particular, hardware platforms, together with multi-processing platforms, have allowed verification algorithms to distribute tasks executions across multiple processors, which generate an increase in performance if compared to single-core solution. However, most verification algorithms still disregard the limitations of the CMOS technology, which limits the increase of the chip’s frequency after it reaches 4 GHz.

\textcolor{blue}{Deve-se citar as referencias usando o comando cite do latex}

\textcolor{Red}{\textcolor{blue}{Descrever brevemente o funcionamento dos algortimos propostos} Here, we exploit the availability of multi-core processors; in particular, SMT-based verification methods applied to the HW-SW partition propose are proposed, then the results are compared using ILP (integral linear programming), GA (generic algorithms) in a multi-core version, and the vZ tool that is a state-of-art optimization tool [32].}

\textcolor{Red}{The ILP algorithm uses intlinprog of the lib "Optimization Toolbox"[35], it's function is to search the lower value that satisfies the problem described in Eq.(1). however, for some types of problems is impossible to use ILP due to the high computational cost, GA is a way to optimize the processing time at the expense of precision, since GA does not guarantee to find the maximum or minimum of the problem. vZ uses optimization functions to rewrite the SMT formulas to find the ideal solution, vZ supports its solution in engines(OptSMT, MaxSMT).}

To the best of our knowledge, this is the first work to use a multi-core SMT-based verification to solve a HW-SW partitioning problem in embedded systems. We implement our ideas with the Efficient SMT-based Bounded Model Checker (ESBMC) tool [14]. As its main contribution, this paper shows that it is possible to take advantage of an SMT-based BMC tool in a multi-core architecture to solve optimization problems.

\textcolor{blue}{Voces precisam citar as secoes usando o comando ref do latex. Para isso, cada secao deve conter um label.}
\textcolor{Red}{This article is organized as follows: Section~\ref{background} gives a background on optimization, optimization with vZ, use of BMC technique with ESBMC, multi-cre ESBMC using OpenMP, ESBMC using optimized techniques of search to multiples cores with OpenMP. Section 3 describes the informal and formal mathematical modeling. Section 4 briefly describes th ILP and GA. The BMC method is based on SMT and is presented in section 5. The section 6 presents the paritioning model using vZ. The section 7 presents the experimental evaluation. The Section 8 discusses about the related works. Section 9 presents the conclusions and the future work.}

@@

\textcolor{blue}{Introduzir uma breve descrição da seção.}

\textcolor{red}{Optimization is the act of obtaining the best result (i.e., the optimal solution) under given circumstances [9]. In the design, construction, and maintenance of any engineering system, engineers have to make many technological and managerial decisions at several stages. The ultimate goal of all such decisions is either to minimize the effort required or to maximize the desired benefit. Because the effort required or the benefit desired in any practical situation can be expressed as a function of certain decision variables, optimization can be defined as the process of finding the conditions that give the maximum or minimum value of a function [9].}
%----------------------------------------------
\subsection{Optimization}
%----------------------------------------------

There is no single method available for solving all optimization problems efficiently [9]. The most known technique is linear programming, which is an method applicable for the solution of problems in which the objective function and the constraints appear as linear functions of the decision variables. A particular case of linear programming is ILP, in which the variables can assume just integer values. Eq. (1) shows a typical linear programming problem, where  and  are vectors or matrixes that describe the constraints
\begin{equation}
  minf^t x \: such that  = 
  \begin{cases}
    A.x \leq b, \\ 
    Aeq.x = beq, \\ 
    x \geq 0.
  \end{cases}
\end{equation}

In some cases, the time to find a solution using ILP is impractical. Even with the use of powerful computers, a problem can take hours belerunning before an optimal solution is reached. If the optimization problem is complex, some heuristics can be used to solve the same problem faster, e.g., those used in the GA [9]. The only drawback is that the found solution may not be the global minimum or maximum.

\textcolor{red}{Tools such as vZ and ESBMC also explores the optimization problem, both tools have its main purpose the precision, problems with a high quantity of data make the time an obstacle. The sections 2.2 and 2.3 shows how vZ and ESBMC uses its techniques to solve the paritioning problems, from those techniques an exact answer and reasonable time is exprected.}

%----------------------------------------------
\subsection{\textcolor{Red}{Optimization with Vz}}
%----------------------------------------------

A SMT solver is a tool that verifies formula's satisfiability based in a set of rules SMT. Z3 is a SMT solver made by Microsoft Research, freely available to the community, that supports a variety of theories. Z3 brings a solver DPLL together with a correspondence machine for quantifiers. Z3 is implemented in C++.
vZ is one of the tools implemented on the Z3 solver, it's base function is to optimize objectives within the logic context of constraints, it stands for the functions that make possible formulate optimized criteria.
vZ includes an incremental version of MaxRes to MaxSAT and a Simplex to solve numbers without defined patterns, in it's core MaxSAT is responsible to the restrictions while OptSMT optimize the linear arithmetic objectives.
Among all progress of Z3 using vZ the functions hightlights: maximize, minimize e assert-soft.

\begin{itemize}
\item{\textbf{maximize(T)}
this instruction informs to the solver that a given variable T should be maximized, if possible maximize real, integer and bit vector values.}
\item{\textbf{minimize(T)}
this instruction informs the solver that a given variable T should be minimized, the accepted types are the same as maximize.}
\item{\textbf{Assert-Soft F :weight n}
the instruction Asser-soft is a restriction to F, if possible add a weight n, the default is 1.}
\end{itemize}

Lets optimize (K+W), with the restrictions in (K \textless 2) and (W-K \textless 1), the expected result of the optimization described in the code below is 2. the model generated by vZ shows that K = 1 and W = 1.

\begin{lstlisting}[caption=Code "SMT" using vZ, label=vZ]
1. (declare-const K Int) 
2. (declare-const W Int)
3. (assert (< K 2)) 
4. (assert (< (- W K) 1))
5. (maximize (+ K W)) 
6. (check-sat)
\end{lstlisting}

Accordingly to [34] the architecture of vZ can be described on the image XXX, initially the formulas and the SMT objectives are converted to problems MaxSAT, if there are many objectives, vZ invokes MaxSAT or OptSAT. To restrictions in real values using “Box”, vZ combines all linear arithmetic and calls the OptSMT. When we use “soft constrains” in the mode “lexicographic” vZ invokes MaxSAT.
\begin{figure}[ht]
	\centering
  \includegraphics[width=0.49\textwidth, height=95px]{Image/vzArch.png} 
	\caption{vZ Architecture}
	\label{fig2}
\end{figure}

Z3 is available for platforms like C, C ++, Java, .NET and Python, it is possible to download the Z3 with vZ in http://z3.codeplex.com. API in python was used to formulate the optimization problem on the vZ tool.

%----------------------------------------------
\subsection{Bounded Model Checking with ESBMC}
%----------------------------------------------

Model checking refers to algorithms for exploring the state space of a transition system to determine if it obeys a speciﬁcation of its intended behavior [3], [4]. These algorithms can perform exhaustive exploration in a highly automatic way and, thus, have attracted much interest in industry. However, model-checking has been held back by the state explosion problem, in which the number of states in a system grows exponentially in the number of system components [10]. Much research has been devoted to mitigate this problem.

Among the recent techniques, there is one that combines model checking with satisﬁability solving. This technique, known as bounded model checking (BMC), does a very fast exploration of the state space, and for some types of problems, it offers large performance improvements over previous approaches, as shown in [10]. In particular, BMC based on Boolean Satisfiability (SAT) has been introduced as a complementary technique to binary decision diagrams for alleviating the state explosion problem.

The basic idea of BMC is to check the negation of a given property at a given depth: given a transition system $ M $, a property $ \phi $, and a bound $ K $, BMC unrolls the system  times and translates it into a verification condition (VC) $ \varphi $  such that $ \varphi $   is satisfiable if and only if $ \phi $ has a counterexample of depth $ K $ or less [10]. To cope with increasing software complexity, SMT solvers can be used as back-ends for solving the generated VCs, as shown in [11], [12], and [13].

According to [14] and [15], SMT-based model checking can be used to verify the single- and multi-threaded software. In [16], ESBMC can also be used to model check C++ software based on SMT solvers. In [5] and [6] it was shown that it is possible to use ESBMC, as an optimization tool.

@@

%----------------------------------------------
\subsection{OpenMP}
%----------------------------------------------

\textcolor{blue}{The OpenMP is a set of directives that augments the C/C++ and Fortran languages. OpenMP uses the fork-join model of parallel execution}

%----------------------------------------------
\section{Mathematical modeling}
%----------------------------------------------

The mathematical modeling was taken from [1], [2].

%----------------------------------------------
\subsection{Informal Model (or Assumptions)}
%----------------------------------------------

The informal model can be described by five characteristics. First, there is only one software context, i.e., there is just one general-purpose processor, and there is only one hardware context. The components of the system must be mapped to either one of these two contexts. Second, the software implementation of a component is associated with a software cost, which is the running time of the component. Third, the hardware implementation of a component has a hardware cost, which can be area, heat dissipation, and energy consumption. Fourth, based on the premise that hardware is significantly faster than software, the running time of the components in hardware is considered as zero. Finally, if two components are mapped to the same context, then there is no overhead of communication between them; otherwise, there is an overhead. The consequence of these assumptions is that scheduling does not need to be addressed in this work. Hardware components do not need scheduling, because the running time is assumed to be zero. Because there is only one processor, software components do not need to be scheduled as well. Therefore, the focus is only on the partitioning problem. That configuration describes a first-generation co-design, where the focus is on bipartitioning [17].

%----------------------------------------------
\subsection{Formal Model}
%----------------------------------------------

The inputs of the problem are: A directed simple graph $ G = (V,E) $, called the task graph of the system, is necessary. The vertices $ V = \{V_1,V_2,\dotso,V_n\} $ represent the nodes that are the components of the system that will be partitioned. The edges $ E $ represent communication between the components. Additionally, each node  $ V_i $ has a cost $ h(V_i) $ (or $ h_i $ of hardware (if implemented in hardware) and a cost $ s(s_i) $ (or $ s_i $ of software (if implemented in software). Finally, $ c(V_i,V_j) $ represents the communication cost between $ V_i $ and $ V_j $ if they are implemented in different contexts (hardware or software).

Based on [1],  is called a hardware-software partition if it is a bipartition of $ V:P = (V_h, V_s) $ , where $ V_h \cup V_s = V $  and $ V_h \cap V_s = 0$  . The crossing $ E_p = \{(V_i,V_j):V_i \in V_s, V_j \in V_h \;or\; v_i \in V_h, v_j \in V_s $ edges are . The hardware cost of $ P $  is given by Eq. (2), and the software cost of $ P $ (i.e., software cost of the nodes and the communication cost) is given by Eq. (3):

\begin{align}
\ H_p &= \Sigma (v_i \;\in\; _Vh) H_i\\
  S_p &= \Sigma (v_i \;\in\; _Vs) S_i + \Sigma (_(v_i,v_j) \in\; _Ep)\; C(V_i, V_j)
\end{align}

Three different optimization and decision problems can be defined. In this paper, the focus is on the case that $ S_0 $ is given, i.e., to find a $ P $ HW-SW partitioning so that $ S_p \leq S_0 $ and $ H_p $ is minimal (system with hard real-time constraints). So, based on Eq. (1) and Eq. (3) the optimization problem’s restrictions can be reformulated as: $ s(1-x) + c|E_x| \leq S_0 $, where $ x $ is the decision variable. Concerning the complexity of this problem, reference [1] demonstrates that it is NP-Hard.

>>>>

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestore
\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[twocolumn]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{color}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{booktabs}
\usepackage{tabu,xcolor,colortbl}
\usepackage{rotating}
\usepackage{amsmath,array}    
\usepackage[utf8]{inputenc} 
\DeclareUnicodeCharacter{FB01}{fi} %Caracter especial
\usepackage{graphicx} %Imagens
\usepackage{listings} % Codes
\usepackage{tabularx} % in the preamble
\usepackage{mathtools} % Serve para as equacoes
\usepackage{cite}

\usepackage{hyphenat}
\usepackage[english]{babel}
\hyphenation{ef-fect-ive consi-de-red demons-trate se-cond veri-fication covera-ge des-cribe reali-zation opera-tor re-pre-sentation methodo-lo-gy di-gi-tal chara-cteristics granu-lar ge-nerally diago-nal para-meters spe-cification re-presentable me-thods ope-rations sche-dule deve-loped des-cribes diffe-rent mo-dels sa-ving ve-rified proper-ties mini-mum mathe-matical tole-rated limi-ted op-tical net-works semi-conduc-tor go-to-pro-grams }

@@

\renewcommand{\baselinestretch}{0.97}

\begin{document}
\title{Applying Multi-Core Model Checking to Hardware-Software Partitioning in Embedded Systems}
\author{Alessandro Trindade\and Hussama Ismail\and Renato Degelo\and Edilson Galv\~ao\and Helder Silva \and Lucas Cordeiro}

\institute{A. Trindade, H. Ismail, R. Degelo, E. Galv\~ao \at
              Graduate Program in Electrical Engineering, Federal University of Amazonas, Brazil \\
%              Tel.: +123-45-678910\\
%              Fax: +123-45-678910\\
              \email{\{iurybessa, hussamaibrahim\}@ufam.edu.br and\\{\{rdegelo, esj.galvao\}@gmail.com}}          %  \\
%             \emph{Present address:} of F. Author  %  if needed
           \and
           H. Silva and L. Cordeiro \at
              Federal University of Amazonas, Brazil
              \email: prhsilva2012@gmail.com lucascordeiro@ufam.edu.br           %  \\
}

@@

\begin{abstract}
We present an alternative approach to solve the hardware (HW) and software (SW) partitioning problem, which uses Bounded Model Checking (BMC) based on Satisfiability Modulo Theories (SMT) in conjunction with a multi-core support using Open MultiProcessing. The multi-core SMT-based BMC approach allows initializing many verification instances based on the number of available processing cores. Each instance checks for a different optimum value until the optimization problem is satisfied. The goal is to show that multi-core model-checking techniques can be effective, in particular cases, to find the optimal solution of the HW-SW partitioning problem using an SMT-based BMC approach. 
We implement and evaluate our algorithms on top of the Efficient SMT-Based Context-Bounded Model Checker (ESBMC). We also compare our approaches to other state-of-the-art optimization tools: Matlab and Z3. Experimental results show that no single optimization tool solves all benchmarks; however, Matlab and Z3 are the most efficient ones to solve HW-SW partitioning problems, although ESBMC had a significant performance improvement from its sequential to parallel version.
\end{abstract}

\keywords{hardware-software co-design \and hardware-software partitioning\and optimization\and model checking\and multi-core\and OpenMP }

%---------------------------------
\section{Introduction}
\label{Introduction}
%---------------------------------

Nowadays, with the strong development of embedded systems, the design phase plays an important role. At early stages, the design is split into separated flows: hardware and software. Consequently, the partitioning decision process, which deals with the decisions upon which parts of the application have to be designed in hardware (HW) and which in software (SW), must be supported by any well-structured methodology. If not, this leads to a number of issues (design flow interruptions, redesigns, and undesired iterations) which affects the overall development process, the quality and the lifecycle of the final system. Starting at the 1990s, intensive research was performed, and several approaches proposed, as shown in ~\cite{Arato2003} and ~\cite{Mann2007}.

In any HW and SW design of complex systems, more time is spent on verification than on construction ~\cite{Baier2008}. Formal methods based on model checking offer great potential to obtain a more effective and faster verification in the design process. Programs may be viewed as mathematical objects with behavior that is, in principle, well determined. This makes it possible to specify programs using mathematical logic, which constitutes the intended (correct) behavior. Then, one can try to give a formal proof or otherwise establish that the program meets its specification ~\cite{Clarke2009}. Research in formal methods has led to the development of very promising verification techniques, which facilitate the early detection of errors. Model-based verification techniques use models that describe the possible system behavior in a mathematically precise and unambiguous manner. The system models are accompanied by algorithms that systematically explore all the states of the system model.

In ~\cite{Trindade2015} and ~\cite{Trindade2014} was shown that it is possible to use Bounded Model Checking (BMC) based on Satisfiability Modulo Theories (SMT) to perform HW-SW partitioning in embedded systems. The present work extends those studies since there is a substantial improvement in terms of the genetic algorithm and the SMT-based verification method, which has been extended with a multi-core architecture. Multi-core processors have been used in all segments of industry to implement high-performance computing ~\cite{Wu2014}. In particular, hardware platforms, together with multi-processing platforms, have allowed verification algorithms to distribute tasks executions across multiple processors, which generate an increase in performance if compared to single-core solution. However, most verification algorithms still disregard the limitations of the CMOS technology, which limits the increase of the chip’s frequency after it reaches 4 GHz.

Here, we exploit the availability of multi-core processors; in particular, SMT-based verification methods are applied to the HW-SW partition problem and experimental results are compared to ILP (integral linear programming), GA (generic algorithms) in a multi-core version, and also the vZ tool, which is a state-of-art optimization tool based on SMT~\cite{Bjorner2015}. The ILP and GA algorithms are implemented with the Optimization Toolbox of Matlab~\cite{OpenMP1998}, while vZ is a built-in tool to the SMT solver Z3 in which its main function is to check the satisfiability of logical formulas based on objective functions. To the best of our knowledge, this is the first work to use a multi-core SMT-based verification to solve a HW-SW partitioning problem in embedded systems. We implement our ideas with the Efficient SMT-based Bounded Model Checker (ESBMC) tool ~\cite{Cordeiro2012}. As its main contribution, this paper shows that it is possible to take advantage of an SMT-based BMC tool in a multi-core architecture to solve optimization problems.

This article is organized as follows: Section~\ref{background} gives a background on optimization techniques, vZ, ESBMC, and OpenMP tools. 
Section~\ref{Mathematical-modeling} describes the informal and formal mathematical modeling. Section~\ref{ILPGA} briefly describes the ILP and GA algorithms, while Section~\ref{Analysis-of-the-partitioning-problem-using-vZ} presents the partitioning model using vZ. The SMT-based BMC method is presented in Section~\ref{Analysis-of-the-partitioning-problem-using-ESBMC}. In Section~\ref{Experimental-Evaluation}, we present the results of our experiments using several embedded systems applications. In Section~\ref{Related-Work}, we discuss the related work and we conclude and describe future work in Section~\ref{Conclusions}.

@@

The HW-SW partitioning problem is typically represented as a set of constraints and an objective function in linear programming. In this section, we describe the linear programming problem and present related tools that are used to model and solve the HW-SW partitioning problem.

%----------------------------------------------
\subsection{Optimization}
\label{Optimization}
%----------------------------------------------

Optimization is the act of obtaining the best result (i.e., the optimal solution) under given circumstances~\cite{Rao2009}. In the design, construction, and maintenance of any engineering system, engineers have to make many technological and managerial decisions at several stages. The ultimate goal of all such decisions is either to minimize the effort required or to maximize the desired benefit. Because the effort required or the benefit desired in any practical situation can be expressed as a function of certain decision variables, optimization can be defined as the process of finding the conditions that give the maximum or minimum value of a function~\cite{Rao2009}.

There is no single method available for solving all optimization problems efficiently~\cite{Rao2009}. The most well-known technique is linear programming, which is an method applicable for the solution of problems in which the objective function and the constraints appear as linear functions of the decision variables. A particular case of linear programming is ILP, in which the variables can assume just integer values. Eq.~\ref{linear-programming-problem} shows a typical linear programming problem, where $A$ and $b$ are vectors or matrixes that describe the constraints
\begin{equation}
\label{linear-programming-problem}
  minf^t x \: such that  = 
  \begin{cases}
    A.x \leq b, \\ 
    Aeq.x = beq, \\ 
    x \geq 0.
  \end{cases}
\end{equation}

In some cases, the time to find a solution using ILP is impractical. Even with the use of powerful computers, a problem can take hours before an optimal solution is reached. If the optimization problem is complex, some heuristics can be used to solve the same problem faster, e.g., those used in the GA~\cite{Rao2009}. The only drawback is that the found solution may not be the global minimum or maximum. Alternatively, tools such as vZ and ESBMC can be used to solve optimization problems so that the global minimum or maximum solution is found. The following sections describe the main features of vZ and ESBMC tools.

%----------------------------------------------
\subsection{Solving Optimization Problems with Vz}
\label{Optimization-with-Vz}
%----------------------------------------------

An SMT solver is a tool that verifies formula satisfiability based on a set of background theories (e.g., linear arithmetic, bitvectors, and arrays). Z3 is built on top of a Davis–Putnam–Logemann–Loveland solver together with auxiliary algorithms for solving quantifiers formulas~\cite{Bjorner2014}. vZ is one of the tools implemented in the SMT solver Z3; its base function is to optimize objective functions, which formulate optimized criteria, within the logical context of constraints.  vZ includes an incremental version of  \textcolor{red}{Maximum Resiliency (MaxRes) ~\cite{Federica2008} to Maximum Satisfiability (MaxSAT)~\cite{NarodytskaN}} and a Simplex to solve numbers without defined patterns. In Vz, MaxSAT is responsible for the restrictions, while OptSMT optimizes linear arithmetic objectives~\cite{Bjorner2015}. Vz provides functions that extend Z3 for solving optimization problems: \textit{maximize}, \textit{minimize}, and \textit{assert-soft}.

\begin{itemize}
\item{\textbf{maximize(T)}
this function informs to the solver that a given variable $T$ should be maximized, which includes real, integer, or bit-vector variables.}
\item{\textbf{minimize(T)}
this function informs the solver that a given variable $T$ should be minimized, the accepted types are the same as maximize function.}
\item{\textbf{Assert-Soft F : weight n}
the function Assert-Soft adds a restriction to $F$, which can also add a weight $n$; the default value is $1$.}
\end{itemize}

As an example, one can optimize $\left(K + W\right)$, with restrictions in $\left(K < 2\right)$ and $\left(W - K < 1\right)$. The expected result of this optimization problem described in the code below is $2$. In fact, the model generated by vZ shows that $K = 1$ and $W = 1$.

\begin{lstlisting}[caption=SMT formula using vZ, label=vZ]
1. (declare-const K Int) 
2. (declare-const W Int)
3. (assert (< K 2)) 
4. (assert (< (- W K) 1))
5. (maximize (+ K W)) 
6. (check-sat)
\end{lstlisting}

The vZ architecture is shown in Fig.~\ref{vZ-Architecture}. Initially, the SMT formula with objectives is converted to $0-1$ constraints, which leads to a Pseudo-Boolean Optimization (PBO)\textcolor{red}{~\cite{Vasco2005}~\cite{Barth1995}}. If there are many objective functions, vZ invokes OptSAT for arithmetic or MaxSAT for soft constraints. \textcolor{red}{for box constraints using real values, Vz combines linear arithmetic objectives and uses only one instance of OptSMT. When we use "soft constrains'' in the mode "lexicographic'' vZ invokes MaxSAT using multiple calls for your engine.} 
%
\begin{figure}[ht]
	\centering
  \includegraphics[width=0.49\textwidth, height=95px]{Image/vzArch.png} 
	\caption{vZ Architecture~\cite{Bjorner2015}}
	\label{vZ-Architecture}
\end{figure}

Z3 is available for platforms in C, C$++$, Java, .NET and Python; it is possible to download Z3 with vZ from~\cite{Z3API}. In this work, the python API is used to formulate the HW-SW partitioning problem in vZ tool.

%----------------------------------------------
\subsection{Bounded Model Checking with ESBMC}
\label{Bounded-Model-Checking-with-ESBMC}
%----------------------------------------------

Model checking refers to algorithms for exploring the state space of a transition system to determine if it obeys a speciﬁcation of its intended behavior ~\cite{Baier2008}, ~\cite{Clarke2009}. These algorithms can perform exhaustive exploration in a highly automatic way and, thus, have attracted much interest in industry. However, model-checking has been held back by the state explosion problem, in which the number of states in a system grows exponentially in the number of system components ~\cite{Biere2009}. Much research has been devoted to mitigate this problem.

Among the recent techniques, there is one that combines model checking with satisﬁability solving. This technique, known as bounded model checking (BMC), does a very fast exploration of the state space, and for some types of problems, it offers large performance improvements over previous approaches, as shown in ~\cite{Biere2009}. In particular, BMC based on Boolean Satisfiability (SAT) has been introduced as a complementary technique to binary decision diagrams for alleviating the state explosion problem.

The basic idea of BMC is to check the negation of a given property at a given depth: given a transition system $ M $, a property $ \phi $, and a bound $ K $, BMC unrolls the system  times and translates it into a verification condition (VC) $ \varphi $  such that $ \varphi $   is satisfiable if and only if $ \phi $ has a counterexample of depth $ K $ or less ~\cite{Biere2009}. To cope with increasing software complexity, SMT solvers can be used as back-ends for solving the generated VCs, as shown in ~\cite{Clarke2004}, ~\cite{Armando2009}, and ~\cite{Ganai2006}.

According to ~\cite{Cordeiro2012} and ~\cite{Cordeiro2011}, SMT-based model checking can be used to verify the single- and multi-threaded software. In ~\cite{Ramalho2013}, ESBMC can also be used to model check C++ software based on SMT solvers. In ~\cite{Trindade2015} and ~\cite{Trindade2014} it was shown that it is possible to use ESBMC, as an optimization tool.

@@

%----------------------------------------------
\subsection{OpenMP}
\label{OpenMP}
%----------------------------------------------
The OpenMP is a set of directives for parallel programming that augments C/C$++$ and Fortran languages. OpenMP supports most processor architectures and operating systems, e.g., Solaris, AIX, HP-UX, Linux, Mac OS X, and Windows. OpenMP uses a portable and very robust model that makes easy for programmers to develop parallel applications for a variety of platforms. In particular, OpenMP uses the fork-join model of parallel execution ~\cite{OpenMP1998}

The most basic directive of OpenMP is the ``\#pragma omp parallel for'', which parallelizes the enclosing loop; a basic OpenMP example is shown below:

\textcolor{blue}{Nao existem outras diretivas? Explicar um pouco mais sobre o OpenMP, em especial, tente descrever as funcoes que usamos na nossa solucao. Foi somente o ``\#pragma omp parallel for''??}

\begin{lstlisting}[caption=OpenMP basic sample]
01.  int i;
02.  #pragma omp parallel for
03.  for (i = 0; i < 10; i++)
04.      a[i] = 2 * i;
\end{lstlisting}

In the above example, the for loop is executed in parallel \textcolor{blue}{Como? Cada posicao do array eh executada em um nucleo diferente?}. Each iteration of the loop is executed in a separated thread.

%----------------------------------------------
\section{Mathematical modeling}
\label{Mathematical-modeling}
%----------------------------------------------

The mathematical modeling of the HW-SW partitioning problem was taken from~\cite{Arato2003,Mann2007}.

%----------------------------------------------
\subsection{Informal Model (or Assumptions)}
\label{Informal-Model-or-Assumptions}
%----------------------------------------------

The informal model can be described by five characteristics. First, there is only one software context, i.e., there is just one general-purpose processor, and there is only one hardware context. The components of the system must be mapped to either one of these two contexts. Second, the software implementation of a component is associated with a software cost, which is the running time of the component. Third, the hardware implementation of a component has a hardware cost, which can be area, heat dissipation, and energy consumption. Fourth, based on the premise that hardware is significantly faster than software, the running time of the components in hardware is considered as zero. Finally, if two components are mapped to the same context, then there is no overhead of communication between them; otherwise, there is an overhead. The consequence of these assumptions is that scheduling does not need to be addressed in this work. Hardware components do not need scheduling, because the running time is assumed to be zero. Because there is only one processor, software components do not need to be scheduled as well. Therefore, the focus is only on the partitioning problem. That configuration describes a first-generation co-design, where the focus is on bipartitioning ~\cite{Teich2012}.

%----------------------------------------------
\subsection{Formal Model}
\label{Formal-Model}
%----------------------------------------------

The inputs of the problem are: a directed simple graph $ G = (V,E) $, called the task graph of the system, is necessary. The vertices $V = \{V_1,V_2,\dotso,V_n\}$ represent the nodes that are the components of the system that will be partitioned. The edges $E$ represent communication between components. Additionally, each node  $V_i$ has a cost $h(V_i)$ (or $h_i$) of hardware if implemented in hardware and a cost $s(s_i)$ (or $ s_i $) of software if implemented in software. Finally, $c(V_i,V_j)$ represents the communication cost between $V_i$ and $V_j$ if they are implemented in different contexts (hardware or software).
